Rule Synthesis for Rule-based Intrusion Detectors

### 2020-06-02 paper/paper.tex
at --> in

Finally, \tname{} ranks plausible rules according to heuristics
obtained by analyzing a database of existing rules. The general
intuition is that new rules should share characteristics with existing
human-written rules. For example, one heuristic function used by
\tname{} leverages the popularity of options for ranking. According to
that heuristic alone, the score of a rule $r$ is proportional to the
popularity of the options declared in
$r$. Figure~\ref{fig:distribution-options} shows the distribution of
options associated with rules from a public
ruleset~\cite{emerging-threats-open} containing
\numrulessuri\ rules. Note that rules with options \CodeIn{content}
and \CodeIn{flow} are the most common. The options \CodeIn{dsize} and
\CodeIn{itype}, used in the golden rule, appear as the 6th and 7th
most popular, respectively. However, when restricted to the set of
five options obtained at the first step, \CodeIn{dsize} and
\CodeIn{itype} are the most popular. At the end of this step, \tname{}
ranks the golden rule second in the list including
\pingscanplausible{} rules. It is worth noting that
\tname\ consistently ranks the golden rule very high in most of the
cases even when the technique produces a large number of plausible
rules (see Table~\ref{table:results}).

@@

For the field strategy, \tname{} parses the input messages looking for
options associated with the data in the fields of
messages. Section~\ref{sec:rules-and-packets} shows the correspondence
between packet fields and some of the rule options from \suri\ and
Section~\ref{sec:active-recon} shows a concrete example where \tname{}
uses this strategy. Although we built the option-field mapping
manually, from the \suri\ documentation, this is a one-time activity.
Unfortunately, for http attacks, the relevant data for the options is
likely at the message payload; not the fields. By analyzing our data
set of rules, we found that this occurs frequently in http attacks,
which are prevalent. For example, \percHttp\ of the \numrulessuri{}
rules from the ruleset we analyzed~\cite{emerging-threats-open}
prescribe defenses against http attacks. In addition,
\percRulesWithContent{} of these rules contain at least one content
option.

@@

\tname{} generates plausible rules while minimizing the seed
rule. Algorithm~\ref{algo1} shows the pseudocode of that
procedure. The algorithm takes as input an over-constrained rule $sr$,
denoting the seed rule produced at the Reverse Engineering step, and
pcap files~\cite{pcap} to reconstruct the benign traffic
($\mathit{bTr}$) and the malicious traffic ($\mathit{mTr}$). The
algorithm produces on output a set of plausible rules. Each of these
rules should capture every instance of the attack (as per
$\mathit{mTr}$) and not capture any message from the benign traffic
(as per $\mathit{bTr}$).

@@

This section evaluates sensitivity of \tname{} to the amount of
examples provided on input. Recall that \tname\ uses one example
attack to generate the seed rule (at the Reverse Engineering step) and
the remaining attacks to evaluate soundness of the rule (at the Rules
Generation step). It is expected that the number and variety of attacks
play a role in the performance of the technique.  We answered this
question in two parts. First, we analyzed the impact of the amount of
positive examples in results. For that, we selected two cases with the
highest number of examples---idqaccess and scripttag---and measured
the position of the golden rule at the rank for an increasing number
\begin{wrapfigure}[9]{r}{0.28\textwidth}  
  \vspace{-2ex}
  \centering
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
    \vspace{-3ex}
    \caption{idqaccess}
  \end{subfigure}%
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
    \vspace{-3ex}        
    \caption{scripttag}
  \end{subfigure}%
    \vspace{-1ex}          
  \caption{\label{fig:impact-number-attacks}Positive examples.}
  \vspace{-2ex}
\end{wrapfigure}
of examples provided to \tname. For each number of variants, we
sampled five sets of that size to provide as input to the technique.
Figure~\ref{fig:impact-number-attacks} shows results for each
attack we selected. The y-axis of each plot denotes the rank of the
golden rule and x-axis denotes the number of attack examples used to
synthesize rules. Each point in the x-axis is associated with the
boxplot of the distribution for the corresponding number of
variants. Results indicate that the number of examples is important,
the choice of examples is important, and a small number of examples
suffices to reach the best results.

>>>>

Finally, \tname{} ranks plausible rules according to heuristics
obtained by analyzing a database of existing rules. The general
intuition is that new rules should share characteristics with existing
human-written rules. For example, one heuristic function used by
\tname{} leverages the popularity of options for ranking. According to
that heuristic alone, the score of a rule $r$ is proportional to the
popularity of the options declared in
$r$. Figure~\ref{fig:distribution-options} shows the distribution of
options associated with rules from a public
ruleset~\cite{emerging-threats-open} containing
\numrulessuri\ rules. Note that rules with options \CodeIn{content}
and \CodeIn{flow} are the most common. The options \CodeIn{dsize} and
\CodeIn{itype}, used in the golden rule, appear as the 6th and 7th
most popular, respectively. However, when restricted to the set of
five options obtained in the first step, \CodeIn{dsize} and
\CodeIn{itype} are the most popular. At the end of this step, \tname{}
ranks the golden rule second in the list including
\pingscanplausible{} rules. It is worth noting that
\tname\ consistently ranks the golden rule very high in most of the
cases even when the technique produces a large number of plausible
rules (see Table~\ref{table:results}).

@@

For the field strategy, \tname{} parses the input messages looking for
options associated with the data in the fields of
messages. Section~\ref{sec:rules-and-packets} shows the correspondence
between packet fields and some of the rule options from \suri\ and
Section~\ref{sec:active-recon} shows a concrete example where \tname{}
uses this strategy. Although we built the option-field mapping
manually, from the \suri\ documentation, this is a one-time activity.
Unfortunately, for http attacks, the relevant data for the options is
likely in the message payload; not the fields. By analyzing our data
set of rules, we found that this occurs frequently in http attacks,
which are prevalent. For example, \percHttp\ of the \numrulessuri{}
rules from the ruleset we analyzed~\cite{emerging-threats-open}
prescribe defenses against http attacks. In addition,
\percRulesWithContent{} of these rules contain at least one content
option.

@@

\tname{} generates plausible rules while minimizing the seed
rule. Algorithm~\ref{algo1} shows the pseudocode of that
procedure. The algorithm takes as input an over-constrained rule $sr$,
denoting the seed rule produced in the Reverse Engineering step, and
pcap files~\cite{pcap} to reconstruct the benign traffic
($\mathit{bTr}$) and the malicious traffic ($\mathit{mTr}$). The
algorithm produces on output a set of plausible rules. Each of these
rules should capture every instance of the attack (as per
$\mathit{mTr}$) and not capture any message from the benign traffic
(as per $\mathit{bTr}$).

@@

This section evaluates sensitivity of \tname{} to the amount of
examples provided on input. Recall that \tname\ uses one example
attack to generate the seed rule (at the Reverse Engineering step) and
the remaining attacks to evaluate soundness of the rule (at the Rules
Generation step). It is expected that the number and variety of attacks
play a role in the performance of the technique.  We answered this
question in two parts. First, we analyzed the impact of the amount of
positive examples in results. For that, we selected two cases with the
highest number of examples---idqaccess and scripttag---and measured
the position of the golden rule in the ranking for an increasing number
\begin{wrapfigure}[9]{r}{0.28\textwidth}  
  \vspace{-2ex}
  \centering
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
    \vspace{-3ex}
    \caption{idqaccess}
  \end{subfigure}%
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
    \vspace{-3ex}        
    \caption{scripttag}
  \end{subfigure}%
    \vspace{-1ex}          
  \caption{\label{fig:impact-number-attacks}Positive examples.}
  \vspace{-2ex}
\end{wrapfigure}
of examples provided to \tname. For each number of variants, we
sampled five sets of that size to provide as input to the technique.
Figure~\ref{fig:impact-number-attacks} shows results for each
attack we selected. The y-axis of each plot denotes the rank of the
golden rule and x-axis denotes the number of attack examples used to
synthesize rules. Each point in the x-axis is associated with the
boxplot of the distribution for the corresponding number of
variants. Results indicate that the number of examples is important,
the choice of examples is important, and a small number of examples
suffices to reach the best results.
